---
title: "COVID-19 Dashboard"
author: "by Joao Lucas D. R. Hilgert"
date: ''
output:
  flexdashboard::flex_dashboard:
    css: bootstrap.css
    orientation: rows
    smooth_scroll: yes
    source_code: embed
    vertical_layout: scroll
runtime: shiny
---

```{r include=FALSE}
#Copyright: Joao Lucas Dziekan Rodrigues Hilgert, Some rights reserved
#theme: cerulean
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r include = FALSE}
library(tidyr)
library(dplyr)
library(readxl)
library(rnaturalearthdata)
library(rnaturalearth)
library(sp)
library(tmap)
library(leaflet)
library(shiny)
library(rgeos)
library(plotly)
library(zoo)
library(flexdashboard)
library(httr)
library(utils)
library(data.table)
library(rvest)
library(jsonlite)
```

```{r, include=FALSE}
OWIDdata <- fread("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/owid-covid-data.csv") %>%
  mutate(death_rate = total_deaths / total_cases *100)

w.sf <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(sovereignt != "Antarctica")

url2<-"https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/master/data/OxCGRT_latest.csv"
GET(url2, write_disk(tf1 <- tempfile(fileext = ".csv")))
OxCGRT <- read.csv(tf1) %>%
  mutate(Date = as.Date(paste(substr(Date, 1, 4), substr(Date, 5, 6), substr(Date, 7, 8), sep = "-"), format = "%Y-%m-%d")) %>%
  tibble()
```

Home
================================================================

Row {data-height=10}
-----------------------------------------------------------------------
 
Row
-----------------------------------------------------------------------

### Home
<p style="font-size:25px;color:red"><b>General information</b></p> 
This graphical report includes various interactive visualizations on the development of COVID-19 as well as the relating policies around the world.<br/> 
<br/>
The report is made of four sections: A country level visualization of the main data, a spatial distribution of data globally, a tool for comparative analysis, and a spatial and comparative analysis of the stringency of policies related to COVID-19. All graphs are interative. Bars, countries and lines are scrollable and clickable.<br/>
<br/>
To download the raw data from Our World in Data used for this report in .XLSX format [click here](https://covid.ourworldindata.org/data/owid-covid-data.xlsx).<br/> 
<br/>
To download the raw data from the Oxford COVID-19 Government Response Tracker (OxCGRT) used for this report in .CSV format [click here](https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/master/data/OxCGRT_latest.csv).<br/> 
<br/>
Stay happy and healthy!<br/>
<br/>
<p style="font-size:22px;color:red"><b>Donate</b></p> 
The objective of this report is purely informative. **However, I encourage those who have the financial means to donate to a [*Red Cross* or *Red Crescent Society*](https://www.ifrc.org/en/what-we-do/where-we-work/), preferably in a highly affected region.**<br/>
You can also donate to *[Doctors without Borders](https://www.msf.org/donate)*. They provide medical assistance to people affected by conflict, epidemics, disasters, or exclusion from healthcare. 
<br/>
<br/>
*Note:* The number of cases is probably **underreported due to asymptomatic infection and lack of testing in certain regions**. This means that the lethality of the virus is probably lower than the numbers currently indicate. Furthermore, **reporting of data may lag**, which means there might be outliers in the daily data, i.e. days with zero cases or conversely multiples of the actual amount. Additionaly, due to organizational contraints, the current numbers may refer to previous dates in certain regions. As of December 14th 2020, the European Centre for Disease Control and Prevention (ECDC) only reports weekly data. That was incompatible with the way this page works with data and creates graphical representation. Therefore, this dashboard transiotioned to the daily data by Our World in Data.
<br/>
<br/>
*Last data update: `r format(max(OWIDdata$date), "%A, %B %d %Y")`*
<br/>
<br/>
*<span style="color:red">Disclaimer:</span>* The original data has not been altered. The data has been compiled, aggregated and plotted. The author assumes no responsibility or liability for any errors or omissions in the content of this site. The information contained in this site is provided on an "as is" basis with no guarantees of completeness, accuracy, usefulness or timeliness.
 
**© 2020 Joao Lucas D. R. Hilgert. Some rights reserved.**

Country level data {data-navmenu="Data Visualization"}
================================================================

Column {.sidebar}
-----------------------------------------------------------------------

```{r echo=FALSE}
selectInput("graph_variable", "Select variable:",
                c("Cases" = "cases",
                  "Deaths" = "deaths"
                  ))

varlist2 <- list(
  "cases" = c("new_cases","total_cases", "new_cases_smoothed"),
  "deaths" = c("new_deaths", "total_deaths", "new_deaths_smoothed")
  )

varlist2_names <- list(
  "cases" = c("Cases","Cumulative cases", "Daily cases"),
  "deaths" = c("Deaths", "Cumulative deaths", "Daily deaths")
  )

selectInput(
  inputId = "country",
  label = "Select country/block:",
  choices = unique(OWIDdata$location),
  selected = "World"
  )

dateRangeInput("daterange1", "Date range:", start = "2020-01-01", end = as.character(max(OWIDdata$date)), min = "2013-12-31",
                              max = max(OWIDdata$date), format = "yyyy-mm-dd", startview = "year",
                              weekstart = 0, language = "en", separator = " to ",
                              autoclose = TRUE)
```

Data: [Our World in Data](https://github.com/owid/covid-19-data/tree/master/public/data)

Row {data-height=10}
-----------------------------------------------------------------------

First row
-----------------------------------------------------------------------
### Development of daily confirmed `r reactive({input$graph_variable})` and 7-day moving average


```{r, echo=FALSE, fig.cap="Source: World Bank"}
plotfiltered <- reactive({
  subset <- OWIDdata
  subset <- subset[subset$date >= input$daterange1[1] & subset$date <= input$daterange1[2], ]
  subset <- subset[subset$location == input$country, ]
  subset <- subset %>%
    mutate(graph1 = eval(sym(varlist2[[input$graph_variable]][1])), graph2 = eval(sym(varlist2[[input$graph_variable]][2])), graph3 = eval(sym(varlist2[[input$graph_variable]][3])))
})

renderPlotly({
  plotdf <- plotfiltered()
  plot_ly(data = plotdf, x = plotdf$date, y = plotdf$graph1, type = 'bar', name = varlist2_names[[input$graph_variable]][1]) %>% 
    add_trace(x = plotdf$date, y = plotdf$graph3, name = "7 day rolling avg", type='scatter', mode='lines') %>%
    layout(yaxis = list(title = varlist2_names[[input$graph_variable]][1]), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = varlist2_names[[input$graph_variable]][3], color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
})
```

Second row
---------------------------------------------
### Reproduction rate (i.e. R number) - stable (7 days) and sensitive (4 days)

```{r echo=FALSE, fig.cap="Source: World Bank"}
renderPlotly({
  plotdf <- plotfiltered()
  plot_ly(data = plotdf, x = plotdf$date, y = plotdf$reproduction_rate, type = 'bar', name = 'Reproduction rate') %>% 
    add_segments(x = min(plotdf$date), xend = max(plotdf$date), y = 1, yend = 1, line=list(color='#E50000'), name = "R = 1") %>%
    layout(yaxis = list(title = 'Reproduction rate'), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = 'Reproduction rate', color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
})
```


Third row
---------------------------------------------
### Development of total confirmed `r reactive({input$graph_variable})`

```{r echo=FALSE, fig.cap="Source: World Bank"}
renderPlotly({
  plotdf <- plotfiltered()
  plot_ly(data = plotdf, x = plotdf$date, y = plotdf$graph2, type = 'bar', name = 'Daily cases') %>% 
    layout(yaxis = list(title = varlist2_names[[input$graph_variable]][2]), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = varlist2_names[[input$graph_variable]][2], color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
})
```

Fourth row
---------------------------------------------
### Development of deaths as percentage of total confirmed cases

```{r echo=FALSE, fig.cap="Source: World Bank"}
renderPlotly({
  plotdf <- plotfiltered()
  plot_ly(data = plotdf, x = plotdf$date, y = plotdf$death_rate, type = 'bar', name = 'Death rate') %>% 
    layout(yaxis = list(title = 'Death rate (in %)'), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = 'Death rate (in %)', color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
})
```

World Map {data-navmenu="Data Visualization"}
================================================================

Column {.sidebar}
-----------------------------------------------------------------------

```{r echo=FALSE}
OWIDdata_date <- data.frame(value = tapply(OWIDdata$location, OWIDdata$date, length)) %>%
  filter(value > 20) %>%
  mutate(dates = as.Date(rownames(.)))

dateInput("date_adjust", label = "Select date:",
            min = min(OWIDdata$date), max = max(OWIDdata_date$dates), value = as.character(max(OWIDdata_date$dates)), startview = "year")

selectInput("map_variable2", "Select focus:",
                c("Cases" = 1,
                  "Deaths" = 2
                  ))

selectInput("map_variable", "Select range:",
                c("Moving Avg (7 day)" = "movAvg",
                  "Single Day" = "daily",
                  "Cumulative Total" = "total"
                  ))

varlist <- list("daily" = c("new_cases", "new_deaths", "new_cases_per_million", "new_deaths_per_million"),
                "total" = c("total_cases", "total_deaths", "total_cases_per_million", "total_deaths_per_million"),
                "movAvg" = c("new_cases_smoothed", "new_deaths_smoothed", "new_cases_smoothed_per_million", "new_deaths_smoothed_per_million")
                )

varlist_names <- list("daily" = c("Daily cases", "Daily deaths", "Daily cases per million", "Daily deaths per million"),
                  "total" = c("Cumulative cases", "Cumulative deaths", "Cumulative cases per million", "Cumulative deaths per million"),
                "movAvg" = c("Cases (7-day rolling avg)", "Deaths (7-day rolling avg)", "Cases (7-day rolling avg) per million", "Deaths (7-day rolling avg) per million")
                  )

selectInput("scale_variable", "Select scale:",
                c("Logarithmic scale" = "log10_pretty",
                  "Linear scale" = "pretty"
                  ))
```

Data: [Our World in Data](https://github.com/owid/covid-19-data/tree/master/public/data)

Row {data-height=10}
-----------------------------------------------------------------------

Column
-----------------------------------------------------------------------
### `r reactive({varlist_names[[input$map_variable]][as.numeric(input$map_variable2)]})`

```{r, echo=FALSE}
tmap_mode("view")

mapFiltered <- reactive({
  OWIDdata.latest <- OWIDdata %>%
    filter(date == input$date_adjust)
  OWIDdata.w <- w.sf %>%
    left_join(OWIDdata.latest, by = c("gu_a3" = "iso_code"))
})

num1 <- reactive({
  num1 <- input$map_variable2
})


renderLeaflet({
  tm <- tm_shape(mapFiltered()) + 
    tm_polygons(varlist[[input$map_variable]][as.numeric(input$map_variable2)], palette = "YlOrBr", id="sovereignt", style = input$scale_variable) +
    tm_basemap(server = "CartoDB.DarkMatterNoLabels")
  tmap_leaflet(tm)
})
```


Second row
---------------------------------------------
### `r reactive({varlist_names[[input$map_variable]][as.numeric(input$map_variable2)+2]})`

```{r echo=FALSE, fig.cap="Source: World Bank"}
renderLeaflet({
  tm <- tm_shape(mapFiltered()) + 
    tm_polygons(varlist[[input$map_variable]][as.numeric(input$map_variable2)+2], palette = "YlOrBr", id="sovereignt", style = input$scale_variable) +
    tm_basemap(server = "CartoDB.DarkMatterNoLabels")
  tmap_leaflet(tm)
})
```


Comparative Graphs {data-navmenu="Data Visualization"}
================================================================

Column {.sidebar}
-----------------------------------------------------------------------

```{r, echo=FALSE}
selectInput("comp_variable", "Select variable:",
                c("Cases" = "cases",
                  "Deaths" = "deaths"
                  ))

varlist3 <- list(
  "cases" = c("new_cases","total_cases", "new_cases_smoothed", "total_cases_per_million", "new_cases_smoothed_per_million"),
  "deaths" = c("new_deaths", "total_deaths", "new_deaths_smoothed", "total_deaths_per_million", "new_deaths_smoothed_per_million")
  )

varlist3_names <- list(
  "cases" = c("Cases","Cumulative Cases", "Cases (7-day Rolling Avg)", "Cases per million", "Cases (7-day rolling avg) per million"),
  "deaths" = c("Deaths", "Cumulative Deaths", "Deaths (7-day Rolling Avg)", "Deaths per million", "Deaths (7-day rolling avg) per million")
  )

dateRangeInput("daterange2", "Date range:", start = as.character(min(OWIDdata$date)), end = as.character(max(OWIDdata$date)), min = "2013-12-31",
                              max = max(OWIDdata$date), format = "yyyy-mm-dd", startview = "year",
                              weekstart = 0, language = "en", separator = " to ",
                              autoclose = TRUE)

selectInput(
  inputId = "country1",
  label = "Select first country/block:",
  choices = unique(OWIDdata$location),
  selected = "World"
  )

selectInput(
  inputId = "country2",
  label = "Select second country:*",
  choices = c("-", unique(OWIDdata$location)),
  selected = "-"
  )

selectInput(
  inputId = "country3",
  label = "Select third country:*",
  choices = c("-", unique(OWIDdata$location)),
  selected = "-"
  )

selectInput(
  inputId = "country4",
  label = "Select fourth country:*",
  choices = c("-", unique(OWIDdata$location)),
  selected = "-"
  )

selectInput(
  inputId = "country5",
  label = "Select fifth country:*",
  choices = c("-", unique(OWIDdata$location)),
  selected = "-"
  )

```

*Optional\

Data: [Our World in Data](https://github.com/owid/covid-19-data/tree/master/public/data)

Row {data-height=10}
-----------------------------------------------------------------------
```{r include=FALSE}
plotfiltered2 <- reactive({
  subset <- OWIDdata
  subset <- subset[subset$date >= input$daterange1[1] & subset$date <= input$daterange1[2], ]
  subset <- subset[subset$location == input$country1 | subset$location == input$country2 | subset$location == input$country3 | subset$location == input$country4 | subset$location == input$country5, ]
  subset <- subset %>%
    mutate(graph1 = eval(sym(varlist3[[input$comp_variable]][1])), graph2 = eval(sym(varlist3[[input$comp_variable]][2])), graph3 = eval(sym(varlist3[[input$comp_variable]][3])), graph4 = eval(sym(varlist3[[input$comp_variable]][4])), graph5 = eval(sym(varlist3[[input$comp_variable]][5])))
})

countryNames <- reactive({
  cNames <- c(input$country1, input$country2, input$country3, input$country4, input$country5)
  cNames[! cNames %in% "-"]
})
```


```{r eval=FALSE, include=FALSE}
Column
-----------------------------------------------------------------------
### Development of daily confirmed `r reactive({input$comp_variable})`

renderPlotly({
  coNames <- countryNames()
  plotdf <- plotfiltered2()
  plot.comp <- plot_ly(data = plotdf, x = plotdf$date[plotdf$location == coNames[1]], y = plotdf$graph1[plotdf$location == coNames[1]], type = 'scatter', mode = "lines", name = coNames[1])
  if (length(coNames) > 1) {
    for (i in 2:length(coNames)) {
      plot.comp <- plot.comp %>% add_trace(x = plotdf$date[plotdf$location == coNames[i]], y = plotdf$graph1[plotdf$location == coNames[i]], name = coNames[i], evaluate = T)
    }
  }
  plot.comp <- plot.comp %>% layout(yaxis = list(title = varlist3_names[[input$comp_variable]][1]), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = varlist3_names[[input$comp_variable]][1], color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
  plot.comp
})
```



```{r eval=FALSE, include=FALSE}
Column
-----------------------------------------------------------------------
### Development of `r reactive({input$comp_variable})` (7-day moving average)
  
renderPlotly({
  coNames <- countryNames()
  plotdf <- plotfiltered2()
  plot.comp <- plot_ly(data = plotdf, x = plotdf$dateRep[plotdf$countriesAndTerritories == coNames[1]], y = plotdf$graph3[plotdf$countriesAndTerritories == coNames[1]], type = 'scatter', mode = "lines", name = coNames[1])
  if (length(coNames) > 1) {
    for (i in 2:length(coNames)) {
      plot.comp <- plot.comp %>% add_trace(x = plotdf$dateRep[plotdf$countriesAndTerritories == coNames[i]], y = plotdf$graph3[plotdf$countriesAndTerritories == coNames[i]], name = coNames[i], evaluate = T)
    }
  }
  plot.comp <- plot.comp %>% layout(yaxis = list(title = varlist3_names[[input$comp_variable]][3]), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = varlist3_names[[input$comp_variable]][3], color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
  plot.comp
})
```

Column
-----------------------------------------------------------------------
### Development of `r reactive({input$comp_variable})` (7-day moving average) per 1 million inhabitants

```{r, echo=FALSE}
renderPlotly({
  coNames <- countryNames()
  plotdf <- plotfiltered2()
  plot.comp <- plot_ly(data = plotdf, x = plotdf$date[plotdf$location == coNames[1]], y = plotdf$graph5[plotdf$location == coNames[1]], type = 'scatter', mode = "lines", name = coNames[1])
  if (length(coNames) > 1) {
    for (i in 2:length(coNames)) {
      plot.comp <- plot.comp %>% add_trace(x = plotdf$date[plotdf$location == coNames[i]], y = plotdf$graph5[plotdf$location == coNames[i]], name = coNames[i], evaluate = T)
    }
  }
  plot.comp <- plot.comp %>% layout(yaxis = list(title = varlist3_names[[input$comp_variable]][5]), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = varlist3_names[[input$comp_variable]][5], color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
  plot.comp
})
```


```{r eval=FALSE, include=FALSE}
Column
-----------------------------------------------------------------------
### Development of total confirmed `r reactive({input$comp_variable})`
  
renderPlotly({
  coNames <- countryNames()
  plotdf <- plotfiltered2()
  plot.comp <- plot_ly(data = plotdf, x = plotdf$date[plotdf$location == coNames[1]], y = plotdf$graph2[plotdf$location == coNames[1]], type = 'scatter', mode = "lines", name = coNames[1])
  if (length(coNames) > 1) {
    for (i in 2:length(coNames)) {
      plot.comp <- plot.comp %>% add_trace(x = plotdf$date[plotdf$location == coNames[i]], y = plotdf$graph2[plotdf$location == coNames[i]], name = coNames[i], evaluate = T)
    }
  }
  plot.comp <- plot.comp %>% layout(yaxis = list(title = varlist3_names[[input$comp_variable]][2]), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = varlist3_names[[input$comp_variable]][2], color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
  plot.comp
})
```


Column
-----------------------------------------------------------------------
### Development of confirmed `r reactive({input$comp_variable})` per 1 million inhabitants
```{r, echo=FALSE}
renderPlotly({
  coNames <- countryNames()
  plotdf <- plotfiltered2()
  plot.comp <- plot_ly(data = plotdf, x = plotdf$date[plotdf$location == coNames[1]], y = plotdf$graph4[plotdf$location == coNames[1]], type = 'scatter', mode = "lines", name = coNames[1])
  if (length(coNames) > 1) {
    for (i in 2:length(coNames)) {
      plot.comp <- plot.comp %>% add_trace(x = plotdf$date[plotdf$location == coNames[i]], y = plotdf$graph4[plotdf$location == coNames[i]], name = coNames[i], evaluate = T)
    }
  }
  plot.comp <- plot.comp %>% layout(yaxis = list(title = varlist3_names[[input$comp_variable]][4]), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = varlist3_names[[input$comp_variable]][4], color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
  plot.comp
})
```

Gov. Response Tracker {data-navmenu="Data Visualization"}
================================================================

Column {.sidebar}
-----------------------------------------------------------------------
```{r, echo=FALSE}
selectInput("ox_variable", "Select index:",
                c("Gov. Response Index" = "GovernmentResponseIndexForDisplay",
                  "Stringency Index" = "StringencyIndexForDisplay",
                  "Containment Health Index" = "ContainmentHealthIndexForDisplay",
                  "Economic Support Index" = "EconomicSupportIndexForDisplay"
                  ))

varlist4_names <- list(
  "GovernmentResponseIndexForDisplay" = "Gov. Response Index",
  "StringencyIndexForDisplay" = "Stringency Index",
  "ContainmentHealthIndexForDisplay" = "Containment Health Index",
  "EconomicSupportIndexForDisplay" = "Economic Support Index"
  )

dateRangeInput("daterange3", "Graph date range:", start = min(OxCGRT$Date), end = as.character(max(OxCGRT$Date)), min = min(OxCGRT$Date),
                              max = max(OxCGRT$Date), format = "yyyy-mm-dd", startview = "year",
                              weekstart = 0, language = "en", separator = " to ",
                              autoclose = TRUE)

dateInput("date_adjust2", label = "Map date:",
            min = min(OxCGRT$Date), max = max(OxCGRT$Date), value = as.character(max(OxCGRT$Date)), startview = "year")

selectInput(
  inputId = "country1.1",
  label = "Select first country/block:",
  choices = unique(OxCGRT$CountryName)
  )


selectInput(
  inputId = "country2.1",
  label = "Select second country:*",
  choices = c("-", unique(OxCGRT$CountryName)),
  selected = "-"
  )

selectInput(
  inputId = "country3.1",
  label = "Select third country:*",
  choices = c("-", unique(OxCGRT$CountryName)),
  selected = "-"
  )

selectInput(
  inputId = "country4.1",
  label = "Select fourth country:*",
  choices = c("-", unique(OxCGRT$CountryName)),
  selected = "-"
  )

selectInput(
  inputId = "country5.1",
  label = "Select fifth country:*",
  choices = c("-", unique(OxCGRT$CountryName)),
  selected = "-"
  )
```

Source: [Hale, Thomas, Noam Angrist, Emily Cameron-Blake, Laura Hallas, Beatriz Kira, Saptarshi Majumdar, Anna Petherick, Toby Phillips, Helen Tatlow, Samuel Webster (2020). Oxford COVID-19 Government Response Tracker, Blavatnik School of Government](https://www.bsg.ox.ac.uk/research/research-projects/coronavirus-government-response-tracker#data).

For more information on the individual indices [click here](https://www.bsg.ox.ac.uk/sites/default/files/2020-09/BSG-WP-2020-032-v7.0.pdf).

Row {data-height=10}
-----------------------------------------------------------------------

Column
-----------------------------------------------------------------------
### Comparative graph for the `r reactive({varlist4_names[[input$ox_variable]]})`
```{r, echo=FALSE}
plotfiltered3 <- reactive({
  subset <- OxCGRT
  subset <- subset[subset$Date >= input$daterange3[1] & subset$Date <= input$daterange3[2], ]
  subset <- subset[subset$CountryName == input$country1.1 | subset$CountryName == input$country2.1 | subset$CountryName == input$country3.1 | subset$CountryName == input$country4.1 | subset$CountryName == input$country5.1, ]
  subset <- subset[subset$RegionName == "", ]
  subset <- subset %>%
    select(CountryName, CountryCode, Date, input$ox_variable) %>%
    rename(ox_var = input$ox_variable) %>%
    group_by(CountryName) %>%
    fill(ox_var, .direction = "down") %>%
    ungroup()
})

#renderDataTable({plotfiltered3()})
countryNames2 <- reactive({
  cNames <- c(input$country1.1, input$country2.1, input$country3.1, input$country4.1, input$country5.1)
  cNames[! cNames %in% "-"]
})

renderPlotly({
  coNames <- countryNames2()
  plotdf <- plotfiltered3()
  plot.comp <- plot_ly(data = plotdf, x = plotdf$Date[plotdf$CountryName == coNames[1]], y = plotdf$ox_var[plotdf$CountryName == coNames[1]], type = 'scatter', mode = "lines", name = coNames[1])
  if (length(coNames) > 1) {
    for (i in 2:length(coNames)) {
      plot.comp <- plot.comp %>% add_trace(x = plotdf$Date[plotdf$CountryName == coNames[i]], y = plotdf$ox_var[plotdf$CountryName == coNames[i]], name = coNames[i])
    }
  }
  plot.comp <- plot.comp %>% layout(yaxis = list(title = varlist4_names[[input$ox_variable]]), barmode = 'stack') %>% 
    layout(plot_bgcolor='#262626', paper_bgcolor='#262626', legend = list(font = list(color = 'white')), xaxis=list(color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'), yaxis=list(title = varlist4_names[[input$ox_variable]], color = 'white', linecolor='white', gridcolor = 'grey', tickfont=list(color='white'), zerolinecolor = 'grey'))
  plot.comp
})
```

Column
-----------------------------------------------------------------------
### Spatial distribution of the `r reactive({varlist4_names[[input$ox_variable]]})`

```{r, echo=FALSE}
mapFiltered2 <- reactive({
  OxCGRT.m <- OxCGRT %>%
    filter(Date == input$date_adjust2) %>%
    select(CountryName, CountryCode, Date, input$ox_variable) %>%
    rename(ox_var = input$ox_variable) %>%
    group_by(CountryName) %>%
    fill(ox_var, .direction = "down") %>%
    ungroup()
  OxCGRT.w <- w.sf %>%
    left_join(OxCGRT.m, by = c("gu_a3" = "CountryCode"))
  OxCGRT.w
})


renderLeaflet({
  tm <- tm_shape(mapFiltered2()) + 
    tm_polygons("ox_var", palette = "YlOrBr", id="sovereignt") +
    tm_basemap(server = "CartoDB.DarkMatterNoLabels")
  tmap_leaflet(tm)
})
```